---
title: "Best coding practices"
author: "Noora Sissala"
date: "2024-06-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# 1 Coding style

## 1.1 Valid variable names

Which of the following are valid/good variable names in R? What is wrong
with the ones that are invalid/bad?

-   var1

-   3way_handshake

-   .password

-   **test**

-   my-matrix-M

-   three.dimensional.array

-   3D.distance

-   .2objects

-   wz3gei92

-   next

-   P

-   Q

-   R

-   S

-   T

-   X

-   is.larger?

## 1.2 Obscure code

```{r}

# The code below works, but can be improved. Do improve it!
myIterAtoR.max <- 5
second_iterator.max<-7
col.NUM= 10
row.cnt =10
fwzy45 <- matrix(rep(1, col.NUM*row.cnt),nrow=row.cnt)
for(haystack in (2-1):col.NUM){
  for(needle in 1:row.cnt) {
if(haystack>=myIterAtoR.max){
fwzy45[haystack, needle]<-NA}
}}

```

```{r}

# Improved code
my_iterator_max <- 5
second_iterator_max <- 7
col_num <- 10
row_num <- 10
mat <- matrix(rep(1, col_num*row_num), nrow=row_num)
for(i in 1:col_num){
  for(j in 1:row_num) {
    if(i >= my_iterator_max){
      mat[i, j] <- NA
    }
  }
}


```

## 1.3 Better formatting

```{r}

# Improve style and formatting of the following code:
simulate_genotype <- function( q, N=100 ) {
  if( length(q)==1 ){
    p <- (1 - q)
    f_gt <- c(p^2, 2*p*q, q^2) # AA, AB, BB
  }else{
    f_gt<-q
  }
  tmp <- sample( c('AA','AB','BB'), size =N, prob=f_gt, replace=T )
  return(tmp)
}

```

```{r}

# Improved code
simulate_genotype <- function(q, N = 100) {
  if(length(q) == 1) {
    p <- (1 - q)
    f_gt <- c(p^2, (2 * p * q), q^2) # AA, AB, BB
  } else {
    f_gt <- q
  }
  tmp <- sample(c('AA', 'AB', 'BB'), 
                size = N, 
                prob = f_gt, 
                replace = TRUE)
  return(tmp)
}


```

## 1.4 Hidden variable

```{r}

# ssign a vector of three last months (abbreviated in English) in a year to a hidden variable my_months
.my_months <- rev(rev(month.abb)[1:3])

```

## 1.5 Pipeline-friendly functions

```{r}

# Modify the function below so that it works with R pipes
my_filter <- function(threshold = 1, x, scalar = 5) {
  x[x >= threshold] <- NA 
  x <- x * scalar
  return(x)
}

```

```{r}

# Modified function
my_filter <- function(x, threshold = 1, scalar = 5) {
  x[x >= threshold] <- NA 
  x * scalar
}

# Test:
c(-5, 5) %>% my_filter()

```

## 1.6 Untidy code

```{r}

# Is the code below correct? Can it be improved?
simulate_phenotype <- function(pop_params, gp_map, gtype) {
  pop_mean <- pop_params[1]
  pop_var <- pop_params[2]
  pheno <- rnorm(n = N, mean = pop_mean, sd = sqrt(pop_var))
  effect <- rep(0, times = length(N))
  for (gt_iter in c('AA', 'AB', 'BB')) {
    effect[gtype == gt_iter] <- rnorm(n = sum(gtype == gt_iter), 
                                      mean = gp_map[gt_iter, 'mean_eff'], 
                                      sd = sqrt(gp_map[gt_iter, 'var_eff']))
  }
  dat <- data.frame(gt = gtype, raw_pheno = pheno, effect = effect, pheno = pheno + effect)
  return(dat)
}

# N is not initialized anywhere 

```

# 2 Structuring code

## 2.1 Computing variance

Write a modular code (function or functions) that computes the sample
standard deviation given a vector of numbers. Decide how to logically
structure the code. Assume there are no built-in R functions for
computing mean and variance available.

```{r}

my_var <- function(x) {
  m <- my_mean(x)
  sd <- my_sd(x, m)
  sd^2
}

my_sd <- function(x, m) {
  sqrt(sum((x-m)^2)/(length(x)-1))
}

my_mean <- function(x) {
  sum(x)/length(x)
}

my_var(c(1,2,3,4,5,6))

```

## 2.2 Writing a wrapper function

You found two functions in two different packages: the randomSampleInt
function that generates a random sample of integer numbers and the
randomSampleLetter function for generating a random sample of letters.
Unfortunately, the functions are called in different ways which you want
to unify in order to use them interchangeably in your code. Write a
wrapper function around the randomSampleLetter that will provide the
same interface to the function as the randomSampleInt. Also, the
randomSampleLetter cannot handle the seed. Can you add this feature to
your wrapper?

```{r}

# Original functions
randomSampleInt <- function(x, verbose, length, seed = 42) {
  if (verbose) {
    print(paste0('Generating random sample of ', length, ' integers using seed ', seed))
  }
  set.seed(seed)
  sampleInt <- sample(x = x, size = length, replace = TRUE)
  return(sampleInt)
} 

randomSampleLetter <- function(N, silent=T, lett) {
  if (!silent) {
    print(paste0('Generating random sample of ', N, ' letters.'))
  }
  sample <- sample(x = lett, size = N, replace = TRUE)
  return(sample)
}

```

```{r}

# Wrapper function for randomSampleLetter
randomSampleLetterWrapper <- function(x, verbose, length, seed = 42) {
  set.seed(seed)
  randomSampleLetter(N = length, silent = !verbose, lett = x)
}


```

```{r}

# Test

## Original function
set.seed(42)
randomSampleLetter(LETTERS, silent = FALSE, N = 5)

## Wrapper function
randomSampleLetterWrapper(LETTERS, verbose = TRUE, length = 5, seed = 42)

```

## 2.3 Customizing plot

Write a wrapper around the graphics::plot function that modifies its
default behavior so that it plots red crosses instead of black points.
Do it in a way that enables the user to modify other function arguments.

```{r}

# Wrappwr function for plot
my_plot <- function(x, col = 'red', pch = 4, ...) {
  
  plot(x = x, col = col, pch = pch, ...)
  
}

```

```{r}

# Test
## Default function
plot(iris$Sepal.Length, xlab = 'Sepal length', ylab = 'Sepal width')

## Wrapper function
my_plot(iris$Sepal.Length, xlab = 'Sepal length', ylab = 'Sepal width')

```

## 2.4 Bonus task: Adding Arguments to a Function

What if you want to pass some additional parameters to a function and,
sadly, the authors forgot to add ... to the list of function arguments.
There is a way out – you can bind extra arguments supplied as alist
structure to the original function arguments retrieved by formals. Try
to fix the function below, so that the call red_plot(1, 1, col = 'red',
pch = 19) will result in points being represented by red circles. Do use
alist and formals and do not edit the red_plot itself!

**alist:** alist handles its arguments as if they described function
arguments. So the values are not evaluated, and tagged arguments with no
value are allowed whereas list simply ignores them. alist is most often
used in conjunction with formals.

**formals:** Get or set the formal arguments of a function.

-   formals(fun = sys.function(sys.parent()), envir = parent.frame())

-   formals(fun, envir = environment(fun)) \<- value

-   For the first form, fun can also be a character string naming the
    function to be manipulated, which is searched for in envir, by
    default from the parent frame. If it is not specified, the function
    calling formals is used.

```{r}

# Original function
red_plot <- function(x, y) { 
  plot(x, y, las=1, cex.axis=.8, ...)
}

# Test

red_plot(1, 1, col = 'red', pch = 19) # Does not work

```

```{r}

# Modify function
formals(red_plot) <- c(formals(red_plot), alist(... = ))

# Test
red_plot(1, 1, col = 'red', pch = 19)


```

## 2.5 Bonus task: Using options

Use options to change the default prompt in R to hello :-) >.

```{r}

# Check which options are available
head(.Options)

```
```{r}

options(prompt = 'hello :-) >')

```


```{r}

# Change back to default
options(prompt = "> ")

```


